---
title: "eyetrackingR Walkthrough"
output:
  html_document:
  toc: true
  theme: united
---
  
Analysis Timestamp:
```{r}
date()
```

## Dependencies

```{r results='hide'}
library(ggplot2, quietly=T)
library(reshape2, quietly=T)
library(lme4, quietly=T)
```

## Load dataset

Ideally, this would be a function in the package (you can embed data in the 'data' folder of a package)

```{r}
load('word-recognition.Rdata')
```

## Load and set data options for **eyetrackingR** library.

```{r results='hide'}
source('./../analyze-eyetracking.R')

# set data options
data_options = set_data_options( 
  trial_column = "Trial",
  time_column = "TimeFromTrialOnset",
  trackloss_column = "TrackLoss",
  aoi_columns = c('Animate','Inanimate'),
  participant_column = "ParticipantName"
)
```

## Verify state of each relevant column.

```{r}
data <- verify_dataset(data, data_options)
```

## Deal with trackloss.

We are going to:

* Treat looks outside of our AOI as if they are trackloss
* Calculate the amount of trackloss in each trial
* Remove trials with over 50% trackloss
* Remove all remaining trackloss samples from our dataset (so that, in each sample, the toddler is looking either to one AOI or the other)

```{r}
# convert non-AOI looks to trackloss
data <- convert_non_aoi_to_trackloss(data, data_options)

# analyze amount of trackloss by subjects and trials
trackloss <- trackloss_analysis(data, data_options, window_start = 15500, window_end = 21000)

# show trackloss
trackloss

data <- clean_by_trackloss(data,
                           data_options, 
                           trial_prop_thresh = .25,
                           window_start = 15500,
                           window_end = 21000)

# remove all trackloss from remaining trials
data <- remove_trackloss(data, data_options)
```

## Subset data to examine "response" window, and create "Target" and "Distractor" columns based on trial

```{r}
# subset to response window
response_window <- subset_by_window(data, data_options, window_start = 15500, window_end = 21000)

# create "Target" and "Distractor" AOI's
response_window$Target <- ifelse(grepl('(Spoon|Cow|Bottle)', response_window$Trial))
```

## Describe overall results, and exclude by number of trials contributed

```{r}
data_summary <- describe_data(response_window, data_options, dv='Animate', c('ParticipantName'))

# exclude participants who contributed <= 2 trials of each type
# (note: if making exclusions here, should re-run analysis so trackloss exclusions are done *after* this)
low_trials <- unique(data_summary[which(data_summary$NumTrials <= 2), 'ParticipantName'])

master_clean <- master_clean[!(master_clean$ParticipantName %in% low_trials$ParticipantName), ]
response_window <- response_window[!(response_window$ParticipantName %in% low_trials$ParticipantName), ]
```

# Familiar Trials

## Primary analyses

```{r}
# subset familiar trials
response_window_familiar <- filter(response_window, TrialType == 'Familiar')

# assign target by trial name
response_window_familiar$Target <- ifelse(response_window_familiar$Trial %in% c('FamiliarBottle','FamiliarCar','FamiliarSpoon'), 'Inanimate', 'Animate')
response_window_familiar$Target <- factor(response_window_familiar$Target)

# aggregate across trials within subjects in time analysis
test_time_familiar <- time_shape(response_window_familiar, data_options, time_bin_size = 100, 
                                 condition_columns = "Target",
                                 aoi = c("Animate")
)

# visualize time results
plot(test_time_familiar, data_options, condition_column = "Target")

# aggregate by subject
response_window_familiar_agg <- window_shape(response_window_familiar, data_options, aoi='Animate', c('Target','Age','MCDI_Total'))

plot(response_window_familiar_agg, data_options, x_axis_column="Age")

subject_means_familiar <- response_window_familiar_agg %>%
  group_by(ParticipantName,Target,Age,MCDI_Total) %>%
  summarise(MeanAnimate = mean(Prop,na.rm=T), N=length(Prop), ArcSin = asin(sqrt(MeanAnimate))) %>%
  ungroup()

ggplot(subject_means_familiar, aes(x=Target, y=MeanAnimate)) + geom_boxplot() + geom_point(position=position_jitter(.1))

# show condition means
condition_means_familiar <- subject_means_familiar %>%
  group_by(Target) %>%
  summarise(Mean=mean(MeanAnimate), SD=sd(MeanAnimate), N=length(MeanAnimate))

condition_means_familiar

# simple t-test between conditions
t.test(ArcSin ~ Target, data=subject_means_familiar, paired=T)

# mixed-effects linear model on subject*trial data
response_window_familiar_agg$TargetC <- ifelse(response_window_familiar_agg$Target == 'Animate', .5, -.5)
response_window_familiar_agg$TargetC <- scale(response_window_familiar_agg$TargetC, center=T, scale=F)

model <- lmer(elog ~ TargetC + (1 + TargetC | Trial) + (1 + TargetC | ParticipantName), data = response_window_familiar_agg, REML = F)
summary(model)
drop1(model,~.,test="Chi")

# growth curve analysis on time series
test_time_familiar$TargetC <- ifelse(test_time_familiar$Target == 'Animate', .5, -.5)
test_time_familiar$TargetC <- scale(test_time_familiar$TargetC, center=T, scale=F)

model <- lmer(elog ~ TargetC*(ot1 + ot2 + ot3 + ot4 + ot5) + (1 + TargetC | Trial) + (1 | ParticipantName), data = test_time_familiar, REML = F)
summary(model)
drop1(model,~.,test="Chi")

# plot GCA model
ggplot(test_time_familiar, aes(x=Time, y=elog, color=Target)) +
  stat_summary(fun.y=mean, geom="point") +
  stat_summary(aes(y=predict(model,test_time_familiar,re.form=NA)), fun.y=mean, geom="line", linetype='solid', size=2)
```

## Onset-contingent analysis

```{r}
# recode AOIs to target & distractor
response_window_familiar$TrialTarget <- ifelse(response_window_familiar$Target == 'Animate', response_window_familiar$Animate, response_window_familiar$Inanimate)
response_window_familiar$TrialDistractor <- ifelse(response_window_familiar$Target == 'Animate', response_window_familiar$Inanimate, response_window_familiar$Animate)

onset_familiar <- onset_shape(response_window_familiar, data_options, onset_time = 15500, fixation_window_length = 100, target_aoi='TrialTarget')

plot(onset_familiar, data_options)

# compare switch times
familiar_switches <- switch_shape(onset_familiar, data_options)

plot(familiar_switches, data_options)

familiar_switches$FirstAOIC <- ifelse(familiar_switches$FirstAOI == 'TrialTarget', .5, -.5)
familiar_switches$FirstAOIC <- scale(familiar_switches$FirstAOIC, center=T, scale=F)

model <- lmer(FirstSwitch ~ FirstAOIC + (1 + FirstAOIC | Trial) + (1 + FirstAOIC | ParticipantName), data=familiar_switches, REML=F)
summary(model)
```

## Bootstrapped splines analysis.

Do a proper within-subjects analysis for the familiar trials.

```{r}
bootstrapped_familiar <- bootstrapped_shape(test_time_familiar, data_options, condition_column = 'Target', within_subj = TRUE, samples = 1000, resolution = 100, alpha = .05, smoother = 'smooth.spline')

plot(bootstrapped_familiar, data_options)

familiar_bootstrapped_intervals <- analyze_bootstraps(bootstrapped_familiar, data_options)

plot(familiar_bootstrapped_intervals, data_options)

familiar_bootstrapped_divergences <- analyze_bootstrapped_divergences(familiar_bootstrapped_intervals, data_options)
familiar_bootstrapped_divergences
```

# Unfamiliar Trials

## Primary analyses

```{r}
# subset unfamiliar trials 
response_window_unfamiliar <- filter(response_window, TrialType == 'Unfamiliar')

# aggregate across trials within subjects in time analysis
test_time_unfamiliar <- time_shape(response_window_unfamiliar, data_options, time_bin_size = 100, 
                                   condition_columns = "Condition",
                                   aoi = c("Animate")
)

# visualize time results
plot(test_time_unfamiliar, data_options, condition_column = "Condition")

# aggregate by subject
response_window_unfamiliar_agg <- window_shape(response_window_unfamiliar, data_options, aoi='Animate', c('Condition','Age','MCDI_Total'))

subject_means_unfamiliar <- response_window_unfamiliar_agg %>%
  group_by(ParticipantName,Condition,Age,MCDI_Total) %>%
  summarise(MeanAnimate = mean(Prop,na.rm=T), N=length(Prop), ArcSin = asin(sqrt(MeanAnimate))) %>%
  ungroup()

ggplot(subject_means_unfamiliar, aes(x=Condition, y=MeanAnimate)) + geom_boxplot() + geom_point(position=position_jitter(.1))

# show condition means
condition_means_unfamiliar <- subject_means_unfamiliar %>%
  group_by(Condition) %>%
  summarise(Mean=mean(MeanAnimate), SD=sd(MeanAnimate), N=length(MeanAnimate))

condition_means_unfamiliar

# simple t-test between conditions
t.test(ArcSin ~ Condition, data=subject_means_unfamiliar, var.equal=T)

# mixed-effects linear model on subject*trial data
response_window_unfamiliar_agg$ConditionC <- ifelse(response_window_unfamiliar_agg$Condition == 'Adjectives', .5, -.5)
response_window_unfamiliar_agg$ConditionC <- scale(response_window_unfamiliar_agg$ConditionC, center=T, scale=F)

model <- lmer(elog ~ ConditionC + (1 + Condition | Trial) + (1 | ParticipantName), data = response_window_unfamiliar_agg, REML = F)
summary(model)
drop1(model,~.,test="Chi")

# growth curve analysis on time series
test_time_unfamiliar$ConditionC <- ifelse(test_time_unfamiliar$Condition == 'Adjectives', .5, -.5)
test_time_unfamiliar$ConditionC <- scale(test_time_unfamiliar$ConditionC, center=T, scale=F)

model <- lmer(elog ~ ConditionC*(ot1 + ot2 + ot3 + ot4 + ot5) + (1 + ConditionC | Trial) + (1 | ParticipantName), data = test_time_unfamiliar, REML = F)
summary(model)
drop1(model,~.,test="Chi")

# plot GCA model
ggplot(test_time_unfamiliar, aes(x=Time, y=elog, color=Condition)) +
  stat_summary(fun.y=mean, geom="point") +
  stat_summary(aes(y=predict(model,test_time_unfamiliar,re.form=NA)), fun.y=mean, geom="line", linetype='solid', size=2)
```

## Zoomed-in wondow

```{r}
# zoom in on active period from previous study (Ferguson, Graf, & Waxman, 2014) and compare conditions
# from 3.15s post-word onset to 4.05s
zoomed_window_unfamiliar <- subset_by_window(response_window_unfamiliar, data_options, window_start = 18650, window_end = 19550)

# aggregate by subject
test_zoomed_window_unfamiliar_agg <- window_shape(zoomed_window_unfamiliar, data_options, aoi='Animate', c('Condition','Age','MCDI_Total'))

subject_means_zoomed_unfamiliar <- test_zoomed_window_unfamiliar_agg %>%
  group_by(ParticipantName,Condition,Age,MCDI_Total) %>%
  summarise(MeanAnimate = mean(Prop,na.rm=T), N=length(Prop), ArcSin = asin(sqrt(MeanAnimate))) %>%
  ungroup()

ggplot(subject_means_zoomed_unfamiliar, aes(x=Condition, y=MeanAnimate)) + geom_boxplot() + geom_point(position=position_jitter(.1))

# show condition means
condition_means_zoomed_unfamiliar <- subject_means_zoomed_unfamiliar %>%
  group_by(Condition) %>%
  summarise(Mean=mean(MeanAnimate), SD=sd(MeanAnimate), N=length(MeanAnimate))

condition_means_zoomed_unfamiliar

# simple t-test between conditions
t.test(ArcSin ~ Condition, data=subject_means_zoomed_unfamiliar, var.equal=T)

# mixed-effects linear model on subject*trial data
test_zoomed_window_unfamiliar_agg$ConditionC <- ifelse(test_zoomed_window_unfamiliar_agg$Condition == 'Adjectives', .5, -.5)
test_zoomed_window_unfamiliar_agg$ConditionC <- scale(test_zoomed_window_unfamiliar_agg$ConditionC, center=T, scale=F)

model <- lmer(elog ~ ConditionC + (1 + Condition | Trial) + (1 | ParticipantName), data = test_zoomed_window_unfamiliar_agg, REML = F)
summary(model)
drop1(model,~.,test="Chi")
```

## Onset-contingent analysis

```{r}
# recode AOIs to target & distractor
onset_unfamiliar <- onset_shape(response_window_unfamiliar, data_options, onset_time = 15500, fixation_window_length = 100, target_aoi='Animate')

plot(onset_unfamiliar, data_options)

# compare switch times
unfamiliar_switches <- switch_shape(onset_unfamiliar, data_options)

plot(unfamiliar_switches, data_options)

unfamiliar_switches$FirstAOIC <- ifelse(unfamiliar_switches$FirstAOI == 'Animate', .5, -.5)
unfamiliar_switches$FirstAOIC <- scale(unfamiliar_switches$FirstAOIC, center=T, scale=F)

model <- lmer(FirstSwitch ~ FirstAOIC + (1 + FirstAOIC | Trial) + (1 + FirstAOIC | ParticipantName), data=unfamiliar_switches, REML=F)
summary(model)
```

## Bootstrapped splines analysis.

Do a between-subjects analysis for the unfamiliar trials.

```{r}
bootstrapped_unfamiliar <- bootstrapped_shape(test_time_unfamiliar, data_options, condition_column = 'Condition', within_subj = FALSE, samples = 1000, resolution = 100, alpha = .05, smoother="smooth.spline")

plot(bootstrapped_unfamiliar, data_options)

unfamiliar_bootstrapped_intervals <- analyze_bootstraps(bootstrapped_unfamiliar, data_options)

plot(unfamiliar_bootstrapped_intervals, data_options)

unfamiliar_bootstrapped_divergences <- analyze_bootstrapped_divergences(unfamiliar_bootstrapped_intervals, data_options)
unfamiliar_bootstrapped_divergences
```

## Individual differences

### MCDI Scores

```{r}
ggplot(subject_means_unfamiliar, aes(x=MCDI_Total, y=MeanAnimate, color=Condition)) +
  geom_point() +
  stat_smooth(method="lm")

with(subject_means_unfamiliar,
     cor.test(MeanAnimate,MCDI_Total))
```

### Age

```{r}
ggplot(subject_means_unfamiliar, aes(x=Age, y=MeanAnimate, color=Condition)) +
  geom_point() +
  stat_smooth(method="lm")

with(subject_means_unfamiliar,
     cor.test(MeanAnimate,Age))
```

### Performance on Familiar Trial

```{r}
# get random effects (centered at 0) for individuals
model <- lmer(elog ~ TargetC + (1 + TargetC | Trial) + (1 + TargetC | ParticipantName), data = response_window_familiar_agg, REML = F)

familiar_accuracy <- data.frame(
  ParticipantName = rownames(ranef(model)$ParticipantName),
  FamiliarAccuracy = ranef(model)$ParticipantName[, 'TargetC']
)

subject_means_unfamiliar <- merge(subject_means_unfamiliar, familiar_accuracy, by='ParticipantName')

ggplot(subject_means_unfamiliar, aes(x=FamiliarAccuracy, y=MeanAnimate, color=Condition)) +
  geom_point() +
  stat_smooth(method="lm")

with(subject_means_unfamiliar,
     cor.test(MeanAnimate,FamiliarAccuracy))
```

### Combine all predictors in a single model...

```{r}
subject_means_unfamiliar$ConditionC <- ifelse(subject_means_unfamiliar$Condition == 'Adjectives', .5, -.5)
subject_means_unfamiliar$ConditionC <- scale(subject_means_unfamiliar$ConditionC, center=T, scale=F)
model <- lm(MeanAnimate ~ ConditionC*(FamiliarAccuracy + Age + MCDI_Total), data=subject_means_unfamiliar)
```